---
title: "ESM 204 HW 3"
author: "Savana Gonzales, Ali Martin, Taylor Medina"
date: '2022-05-04'
output: 
  html_document:
      code_folding: hide
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)

options(scipen = 999)
```

# Setup 
```{r}
# read in the data
data <- read_csv(here('data/HW3_data.csv')) %>% 
  select(!1) 

# perform a linear fit
low.model <- lm(Price_Cents ~ Q_Low_Kwh, data)
high.model <- lm(Price_Cents ~ Q_High_Kwh, data)
```

Functions that will be used throughout the assignment
```{r}
# convert price to demand
demand <- function(price, model) {
  q <- (price - model$coefficients[[1]]) / model$coefficients[[2]]
  q <- ifelse(q < 0, 0, q)
  return(q)
}

# calculate aggregate demand
demand_agg <- function(price, tax) {
  new_price <- price + tax
  q <- demand(new_price, low.model) + demand(price, high.model)
  return(q)
}

# calculate consumer surplus
cs <- function(price, model) {
  q <- demand(price, model)
  cs <- 0.5 * (model$coefficients[[1]] - price) * q
  return(cs)
}

# aggregate consumer surplus
cs_agg <- function(price){
  cs <- cs(price, high.model) + cs(price, low.model)
  return(cs)
}

# producer surplus
ps <- function(price, q_eq){
  ps <- 0.5 * 10 * q_eq
  
  return(ps)
}

eq_p_q <- function(m1, m2, b1, b2){
  q_star <- (b2 - b1) / (m1 - m2)
  p_star <- m1 * q_star + b1
  vector <- c(q_star, p_star)
  return(vector)
}
```

Important values:

Social Cost of Carbon (SCC): $51/tonne

Price/kWh: $0.10

## 1

```{r}
# function that converts quantity in lbs to metric tons and returns mec
mec <- function(q, p){
  q_ton <- q / 2204.62
  p_kwh <- q_ton * p
  return(p_kwh)
}

mec <- round(mec(0.85, 51), 4) * 100

mec
```

# 2

```{r}
price <- seq(0, 40, length.out = 100)
agg_demand <- demand_agg(price, 0) %>% 
  as.data.frame() %>% 
  cbind(price) %>% 
  mutate(demand.low = demand(price, low.model),
         demand.high = demand(price, high.model)) %>% 
  pivot_longer(c(., demand.low, demand.high),
               names_to = 'demand',
               values_to = 'quantity_d') 

agg_demand[agg_demand == 0] <- NA

# lm for agg demand 
agg_demand.lm <- agg_demand %>% 
  filter(quantity_d > 30000, demand == '.')

demand.lm <- lm(price ~ quantity_d, data = agg_demand.lm)

# calculating equilibrium quantity
q_eq <- (10 - demand.lm$coefficients[1]) / demand.lm$coefficients[2]

# slope of supply curve
supply <- 10 / q_eq

ggplot(agg_demand) +
  geom_line(aes(x = quantity_d, y = price, color = demand)) +
  geom_abline(slope = supply) +
  geom_hline(yintercept = mec) +
  theme_minimal() +
  labs(x = "Quantity Demanded (kwh)", y = "Price")
```

```{r}
cs_low <- cs(10, low.model)
cs_high <- cs(10, high.model)
cs_aggr <- cs_agg(10)
p_s <- ps(10, q_eq)
env_cost <- q_eq * mec
```

- The consumer surplus of the low consumers is $`r round(cs_low,2)`.

- The consumer surplus of the high consumers is $`r round(cs_high, 2)`.

- The total benefit is $`r round(cs_aggr, 2)`.

- The producer surplus is $`r round(p_s, 2)`.

- The environmental cost under the status quo is $`r round(env_cost, 2)`.


## 3

```{r}
cs_high_perc <- cs_high/cs_aggr *100
cs_low_perc <- cs_low/cs_aggr *100
```

The high income consumers receive 84.6% of the total consumer benefit, while the low income consumers receive 15.3% of the total consumer benefit 


# 4

a)
```{r}
# make new demand curve under the tax
new_agg_demand <- demand_agg(price, mec) %>%
  as.data.frame() %>% 
  cbind(price) 

new_agg_demand[new_agg_demand == 0] <- NA
  
ggplot(new_agg_demand) +
  geom_line(aes(x = ., y = price)) +
  geom_abline(slope = supply) +
  theme_minimal()

new_aggd.lm <- new_agg_demand %>% 
  filter(. > 250000)

lm.new_demand <- lm(price ~ ., data = new_aggd.lm)

new_eq <- eq_p_q(supply, lm.new_demand$coefficients[2], 0, lm.new_demand$coefficients[1])



```

b)

c)

d)
